Robust data management for high through-put light sheet microscopy of whole mouse brains

### 2015-02-04 latex-style-files-spie-manuscript/manuscript_merged.tex
first commit after Leo - no light sheet comic - change of order: light path first, sample stuff dopo - combined panel for chamber and sample mounting - alignment at bottom in supplementary information

\title{Working title: the most amaz-zing SPIM and how we built it.}

\author{M. Caroline M\"{u}llenbroich,\supscr{a,b} Ludovico Silvestri, \supscr{a,c} Leonardo Onofri,\supscr{e} Irene Costantini,\supscr{a} Marcel Van t'Hoff\supscr{f} Leonardo Sacconi,\supscr{a,c} Giulio Iannello,\supscr{e} Francesco S. Pavone,\supscr{a,b,c,d}  }

@@

The highly ambitious project of mapping and understanding each and every neuronal connection in the whole brain has been moved from the realm of wishful longing to feasible reality by the recent advent of light sheet fluorescent microscopy (LSFM). With this technique 3D data sets can be acquired with a resolution that is high enough to identify neurons and their dendritic, axonal and spine features in time scales which are no longer the bottle neck of high-throughput acquisition. In LSFM, the sample is illuminated with a thin sheet of light confined into the focal plane of the detection objective, which collects the fluorescence emission along an axis perpendicular to the illumination plane\cite{Huisken2009}. This technique drastically reduces the imaging acquisition time by recording millions of pixels in parallel and affords optical sectioning by operating fluorescence excitation and detection on separate, perpendicularly oriented paths where the excitation light sheet and the detection focal plane overlap (Figure\ref{fig:Panel0},A). Consequently fluorophores outside the light sheet are neither bleached nor contribute blurring out-of-focus noise. Consequently LSFM reduces phototoxicity and photobleaching while achieving excellent resolution at high penetration depths; however, it requires the sample to be transparent.

@@

\begin{figure}
   \begin{center}
   \begin{tabular}{c}
   \includegraphics[width=\textwidth]{PanelSampel.eps}
   \end{tabular}
   \end{center}
   \caption{\label{fig:Panel0} (A) Schematic of light sheet microscopy. Excitation (blue) and fluorescence detection (green) occur on independent, perpendicular axes. (B) Mounting of a clarified, fluorescently labelled brain. The brain is glued onto a coverslip and inserted into an adapter that slides into a Teflon cylinder.} 
   \end{figure}

\begin{figure}
   \begin{center}
   \begin{tabular}{c}
   \includegraphics[width=\textwidth]{NuPanel1.eps}
   \end{tabular}
   \end{center}
   \caption{\label{fig:Panel1} Schematic of light sheet microscopy. (A) Fluorescence excitation (along x axis) and detection (along z axis) are operated on independent, perpendicular light paths where the excitation light sheet and the detection focal plane overlap. The custom-made sample chamber is assembled with silicone bellows and seal rings. The clarified, fluorescently-labelled brain is mounted on a Teflon cylinder in the centre of the watertight chamber (C) and can be translated and rotated freely with piezo motors (D).} 
   \end{figure}

To acquire whole brain data sets, we fix a clarified and fluorescently-labelled mouse brain with super glue to a coverslip (Figure\ref{fig:Panel0},B). The brain is oriented along its axial orientation with the olfactory bulbs at the top and the cerebellum at the bottom of the coverslip. The coverslip is slid into a bottom adaptor and tightened with a plastic-capped grub screw. The bottom adaptor is inserted into a Teflon cylinder with the coverslip being positioned on the far side of the detection objective. Three different slits in the adapter correspond to varying distances to the detection objective and give variability in sample thickness for example to allow also for the mounting of rat brain hemispheres.

We designed a cubic, water-tight sample chamber (Figure\ref{fig:Panel1},A) that allows access from all six sides while maintaining the 3D integrity of large, clarified and fluorescently-labelled mouse brains. The sample chamber is tightly bolted  to the optical breadboard while soft connections using silicone bellows allow for adjustive movements of the objectives and free 3D motion of the motor stages (Figure\ref{fig:Panel1},B). All connections are sealed with rubber rings and silicone caulk and additionally tightened with cable binders. In this way the objectives can be refocused and realigned without compromising the watertight seal of the chamber (Figure\ref{fig:Panel1},C). The clarified brains are imaged immersed in clearing solution composed of 63\% TDE in Phosphate buffered saline (PBS) and a refractive index of 1.45. To fill the entire volume of the sample chamber approximately 400ml clearing solution are needed. A motorized x-, y-, z-, $\theta$-stage (Physik Instrumente, see Table\ref{tab:optomechanics}) allows free 3D motion and rotation of the Teflon cylinder which reaches into the centre of the chamber (Figure\ref{fig:Panel1},D). Illumination and detection axes are horizontal and the sampler rotation occurs around the vertical axis.

@@

		\begin{figure}
   \begin{center}
   \begin{tabular}{c}
   \includegraphics[width=\textwidth]{Panel2.eps}
   \end{tabular}
   \end{center}
   \caption{\label{fig:excitation} (A) Topview of the excitation path. The galvo scanners are mounted above periscopes. LP: long-pass filter, I: iris, AOTF: acousto-optical tunable filter, LM: laser modulator, PBS: polarisation beam splitter, ABCD: flip mirrors. Inset: detection. (B) Oblique view of the microscope. A custom-made breadboard serves to mount the sample chamber and objectives at an elevated height and features two circular holes at the edges for the periscopes and a large central cut out for the translation stages. A second breadboard is used for the camera.} 
   \end{figure}

@@

The scan mirror surface is re-imaged with a telescope (f50, f100) onto the back aperture of a long working distance, low magnification objective (Nikon, 10x 0.3NA WD 17.5mm). The two excitation objectives are designed for air immersion but are immersed into the clearing solution within the sample chamber. A coverslip glued to the front housing edge serves the dual purpose of maintaining the first diffractive surface between the front optical element and the air and to protect the front lens elements from the organic solvent. The light sheet is generated digitally\cite{Keller2008a,Keller2008b} by scanning the excitation beam across the focal plane of the detection objective. %This generates perfectly incoherent illumination resulting in fewer artefacts.%cite{StelzerNatMeth2014}.

\subsection{Detection}
		%objective, camera, rolling shutter. Use references\cite{Silvestri2012} to old SPIM.
		%exposure time of single line affects how many are exposed at same time, confocality by synchronisation- not exactly

In it's simplest form, the fluorescence detection unit of a light sheet microscope is a conventional wide-field microscope: an objective lens, a fluorescent filter and a tube lens for an image on a wide-field detector. Fluorescence is collected with an objective that is specifically designed for immersion in clearing solutions. The objective is equipped with a correction collar allowing for immersion in media with refractive indices ranging from 1.41 to 1.52 (Olympus, XLSLPLN25XGMP). To image the whole volume of interest the objective has a relatively low magnification of 25x and a large working distance of 8mm yet a high numerical aperture (1.0NA) affords high resolution imaging. A tube lens of 200mm creates the primary image on a sCMOS camera (Orca Flash4.0, Hamamatsu) with a chip of over 4 megapixels. In contrast to CCD sensors each pixel of a sCMOS camera sensor has a separate photodetector and amplification unit which means that sCMOS pixels can be read out independently from each other. In the rolling shutter data acquisition mode only a subset of adjacent horizontal pixel lines is simultaneously exposed and this active detection region is moved across the image sensor. With the delay between the exposure of two adjacent lines set to the minimum time required to read out a single line ($9.7\micro s$, see Table\ref{tab:resolution}) the entire frame of 2048 horizontal lines is activated within 19.86ms. Setting the line exposure time to 3ms results in approx. 300 lines being simultaneously exposed. This detection area is swept from the top to the bottom of the chip and acts like a moving, virtual confocal slit corresponding to a width of $80\micro m$ in the sample space. With those values the camera acquires images at a frame rate of 44Hz.

@@

@@

\section{Alignment}

%\section{Software Management}
%
	%\begin{figure}
   %\begin{center}
   %\begin{tabular}{c}
   %\includegraphics[width=\textwidth]{connectivity.eps}
   %\end{tabular}
   %\end{center}
   %\caption{\label{fig:connectivity} Connectivity scheme} 
   %\end{figure}
%
%%Here we describe all things software control and communication, components and their interaction. how is the rolling shutter timed with galvos, triggers and clocks etc.
%
%
%%Based on a modular scheme (OOP) using Murmex development kit
%%Separate building blocks control different components of the microscope
%%Hierarchical structure coordinates complex tasks
%%Communication takes place through RabbitMQ server (open source message-oriented middleware) using the LabbitMQ library
%
%The microscope was designed and developed by optical technology specialists yet needs to be operational in a multi-user environment. The greatest challenge therefore consists in making the use of the microscope easy and intuitive such that no formidable expertise is needed when it is operated by interdisciplinary researchers. Our experimental research is based on the acquisition and real-time processing of various data streams. In a given experiment, data may be acquired from multiple, spatially separated sources, and individual processing steps can be computationally intensive and may need to be distributed over several computers to be performed at run-time. For those reasons we chose to write our software in LabVIEW (National Instruments), as this language is particularly well suited for hardware control and data acquisition through pre-developed methods.
%
%Another important aspect of software design is avoiding redundancy when scaling up the software, for example, when a second or third microscope with basic similarity yet slightly specialised equipment is being designed in the same laboratory. Writing well structured, scalable and easily maintainable scientific software in a group of researchers is an often recurring challenge. Object oriented programming is a programming paradigm whose main aspects are encapsulation, inheritance and polymorphism and a useful tool to address these challenges. 
%
%Encapsulation refers to the concept that the entirety of software needed to say operate a microscope is broken up into self-contained modules. Those modules are user-defined data types and referred to as ``classes" where one instance of a class is called an ``object". For example a class could be the basic code to operate an  abstract laser while 5 instances of this class would be the necessary adjustments to operate 5 real world lasers from different manufacturers. This modular programming style keeps large amounts of software routines separate and less prone to conflict with each other.
%
%Classes are arranged in hierarchical structures and inheritance, the act of passing on knowledge from a ``parent" to a ``child" within a class, reduces the amount of programming required when increasingly more complex functions have to be added to the system. For example when specialised behaviour is added to the bottom of the hierarchy, only code regarding this very specific task has to be added while every other aspect is inherited from above. For example, if the code for a motor stage has already been implemented, a rotation stage can be interpreted as the child of a translation stage where the travel range is infinite. This increases code re-usability and saves time.
%
%In this way, the microscope software is decomposed into the communication and interaction of different objects through methods, bits of code, which act on each object's private data.
%
%The Murmex software development kit is based on the OOP paradigm and helps users to collaborate on software development by letting them independently create their respective contributions to the software in the form of new objects or methods within a standardized programming scheme. 
%
%\subsection{RabbitMQ}
%Murmex uses the library LabbitMQ, which is a wrapper of RabbitMQ, a message-oriented middle ware facilitating message shipment and receival. These layers of abstraction are mostly hidden for the developers.
%The RabbitMQ server (message broker) needs to be installed either locally or somewhere in the network. The message-based communication allows for buffered, asynchronous and reliable message communication between several software components. Messages are routed by the broker to the specific software component by its ID. In order to access the functionality of RabbitMQ, the application programming interface (API) \href{http://sine.ni.com/nips/cds/view/p/lang/en/nid/211065}{LabbitMQ} was used, which uses .NET calls for sending and receiving messages to and from the RabbitMQ broker. Messages can be sent to a single recipient by its ID or sent to all available components.
%
%\subsection{Murmex}
%The \href{http://sine.ni.com/nips/cds/view/p/lang/en/nid/212895}{Murmex library} is a software development kit for creating distributed finite state machines in LabVIEW (2012 and newer). A finite state machine resembles the Strategy pattern [Design Patterns: Elements of Reusable Object-Oriented Software, E. Gamma, R. Helm, R. Johnson, J. Vlissides, 1994], which lets the software of one component proceed from one particular state to another based on message it receives from itself or other components (for example to execute Initialise, Start, Proceed, Stop, Terminate). In each state it executes the code for that state only, but flexible execution patterns are possible. It thus allows for predictable behaviour, without knowing the exact code. Murmex has twelve pre-defined common states (e.g. Initialise, Start, Stop, Terminate) and a generic state for implementing custom states. A software component that represents one, yet large, task, often uses only those previously four mentioned states. Examples for such a software component are the implementation of a laser, a translation stage and a camera. Such a component would implement in the Initialise method the initialisation of the communication port to the device, start activity in the Start method, stop activity in the Stop method and close the communication and free up memory in the Terminate. 
%While each component is an autonomous program, variables on which each component might depend, have to be shared among the components. The Update method is used for this purpose, which allows components to update itself with new parameters and automatically forward these parameters to down- or upstream components by using a mapping script $(e.g. datastorage.total_movie_size=xres*yres*number_of_frames, where datastorage is the ID of the recipient, total_movie_size is the name of the parameter to be updated and xres, yres and number_of_frames are the names of parameters of the sender)$.
%Typically, a Murmex component streams data only to components listed as observers in its configuration file. Using this concept, components can be configured to form arbitrarily complex signal processing pipelines or networks. Once a component has been considered to be stable by the developer, an executable can be built. An important component named ServiceManager (also based on Murmex) can communicate to other ServiceManagers over the network or internet and can start remotely executables. This allows the user to start up a sophisticated configuration with multiple components as executable on different computers, but steered from one computer.
%
%
%
%
%\section{Data Management}
	%\subsection{Pipeline}
	%
	%\begin{figure}
   %\begin{center}
   %\begin{tabular}{c}
   %\includegraphics[width=\textwidth]{DataFlow.eps}
   %\end{tabular}
   %\end{center}
   %\caption{\label{fig:DataFlow} Scheme of the data flow. (A) Data streaming on the microscope. The data is written through the CameraLink connection onto the SSD. From there the data can be transferred via a 10Gb connection to the NAS or to the supercomputing centre CINECA in Bologna. SSD: solid-state disk, RAID: redundant array of independent disks, NAS: network-attached storage. (B) Data production rate versus data conversion rate showing the maximum acquisition time and tomography size feasible before the SSD are filled up.} 
   %\end{figure}
	%
%%\begin{landscape}
%\begin{table}%[t!]
	%\centering
		%\caption[Data production]{Data production rates with two python scripts running in parallel.\label{tab:dataproduction}}
		%\begin{tabular}{lrl}
		%frame rate																	& 44			& Hz	\\
		%size of 1 frame															& 8				& MB	\\
		%frames per stack														& 3900		& 		\\
		%size of original stack											& 30.4		& GB	\\
		%size of converted stack											& 3				& GB	\\
		%stack acquisition time											& 88.6		& s		\\
		%stack conversion time												& 140			& s		\\
		%time ratio conversion/ acquisition					& 1.6			& 		\\
		%data production rate												& +1.2		&	TB/h\\
		%data conversion rate												& -0.7		&	TB/h\\
		%net rate																		& +0.5		&	TB/h\\
		%SSD RAID																		& 10.9		& TB	\\
		%max acquisition time to fill up SSDs				& 21			& h		\\
		%max tomo size to fill up SSDs								& 25.4		& TB	\\
		%\end{tabular}
%\end{table}
%%\end{landscape}
%The unprecedented amount of data routinely produced by light sheet microscopy presents challenges to the traditional way we interact with our data. Having far passed the point of taking your data home with you on a hard drive at the end of the experiment, data management in light sheet microscopy requires a professional approach to centralised network architecture. The following paragraphs give an idea of the considerations one has to give to the big data volume problem. The camera used in our light sheet microscope has a dedicated work station with a CameraLink card and 10.9TB of SSD space in RAID 0 configuration. This finite storage capacity is further expanded by a network-attached storage (NAS) of 32.4TB (Figure\ref{fig:DataFlow}). The SSDs and the NAS are connected via a 10Gb bandwidth connection using a 10Gb switch and a 10Gb optical fibre connects our lab to the supercomputing centre CINECA in Bologna. Mention PICO system/storage analysis capacity there? Mention graphical processing unit and cores/processors/RAM of work station?
	%
%With the size of a single image of 8MB and a stack made up of say 3900 frames a single stack amounts to 30.4GB of disc space (see Table\ref{tab:dataproduction} for an overview of the values). Based on a frame rate of 44Hz each stack takes approximately 1.5 minutes to acquire and the SSD RAID fills up with a data writing rate of 1.2TB/h. If no further measures are taken the SSD RAID of 10.9TB would be filled up in just 9h. A custom compression script written in Python serves the triple purpose of translating the images from the proprietary Hamamatsu file format to 3D tif, to transfer the data from the SSD RAID to a network attached storage (NAS) and to reduce the raw data size by first, digitally downsampling the image resolution and second, applying a lossless compression to the tif images. This leads to a overall reduction of data size by a factor of ten, each stack now taking up only 3GB of disc space, albeit at a much slower conversion rate (-0.7TB/h) than the rate at which data is written to the SSD thus representing the most serious bottleneck to our data production pipeline. The SSD RAID fills up at a rate of 0.5TB/h which means the total acquisition time is limited to less than 24h and a total tomo size of just over 25TB. After this time acquisition has to be halted until the slower conversion process has caught up and freed the SSD RAID before data acquisition can start again. A whole mouse brain tomography would take 3 days of uninterrupted measuring yet with the necessary breaks for data transfer this time is easily doubled to tripled. 
%
	%\subsection{Stitching and visualisation}
	%
%Given the limited field of view of the microscope, the acquisition of macroscopic specimens required many parallel image stacks to cover all the volume. Thus, in order to achieve a 3D image of the whole specimen from raw data, the Terastitcher\cite{Bria2012} has been recently proposed, i.e. a stitching tool capable to deal with teravoxel-sized images. However, the Terastitcher does not support input data acquired through the serial sectioning procedure, which leads to a specimen partitioned in different layers. Furthermore, only single channel images can be processed. For these reasons, we extended the Terastitcher functionalities introducing the two following additional features: i) stitching of a specimen partitioned in a number of overlapping layers for the hippocampus reconstruction and, ii) coping with images containing more than one channel for the human brain tomography.

>>>>

\title{Robust data management for high through-put light sheet microscopy of whole mouse brains}

\author{M. Caroline M\"{u}llenbroich,\supscr{a,b} Ludovico Silvestri, \supscr{a,c} Leonardo Onofri,\supscr{e} Irene Costantini,\supscr{a} Marcel Van t'Hoff,\supscr{f} Leonardo Sacconi,\supscr{a,c} Giulio Iannello,\supscr{e} Francesco S. Pavone,\supscr{a,b,c,d}  }

@@

The highly ambitious project of mapping and understanding each and every neuronal connection in the whole brain has been moved from the realm of wishful longing to feasible reality by the recent advent of light sheet fluorescent microscopy (LSFM). With this technique 3D data sets can be acquired with a resolution that is high enough to identify neurons and their dendritic, axonal and spine features in time scales which are no longer the bottle neck of high-throughput acquisition. In LSFM, the sample is illuminated with a thin sheet of light confined into the focal plane of the detection objective, which collects the fluorescence emission along an axis perpendicular to the illumination plane\cite{Huisken2009}. This technique drastically reduces the imaging acquisition time by recording millions of pixels in parallel and affords optical sectioning by operating fluorescence excitation and detection on separate, perpendicularly oriented paths where the excitation light sheet and the detection focal plane overlap. Consequently fluorophores outside the light sheet are neither bleached nor contribute blurring out-of-focus noise. Consequently LSFM reduces phototoxicity and photobleaching while achieving excellent resolution at high penetration depths; however, it requires the sample to be transparent.

@@

%\begin{figure}
   %\begin{center}
   %\begin{tabular}{c}
   %\includegraphics[width=\textwidth]{PanelSampel.eps}
   %\end{tabular}
   %\end{center}
   %\caption{\label{fig:Panel0} (A) Schematic of light sheet microscopy. The sample is illuminated with a planar light sheet (blue) perpendicular to fluorescence detection (green). (B) Mounting of a clarified, fluorescently labelled brain. The brain is glued onto a coverslip and inserted into an adapter that slides into a Teflon cylinder.} 
   %\end{figure}

@@

		\begin{figure}
   \begin{center}
   \begin{tabular}{c}
   \includegraphics[width=\textwidth]{frame1.eps}
   \end{tabular}
   \end{center}
   \caption{\label{fig:excitation} (A) Topview of the excitation path. The galvo scanners are mounted above periscopes. LP: long-pass filter, I: iris, AOTF: acousto-optical tunable filter, LM: laser modulator, PBS: polarisation beam splitter, ABCD: flip mirrors. Inset: detection. (B) Oblique view of the microscope. A custom-made breadboard serves to mount the sample chamber and objectives at an elevated height and features two circular holes at the edges for the periscopes and a large central cut out for the translation stages. A second breadboard is used for the camera.} 
   \end{figure}

@@

The scan mirror surface is re-imaged with a telescope (f50, f100) onto the back aperture of a long working distance, low magnification objective (Nikon, 10x 0.3NA WD 17.5mm). The two excitation objectives are designed for air immersion but are immersed into the clearing solution within the sample chamber. A coverslip glued to the front housing edge serves the dual purpose of maintaining the first diffractive surface between the front optical element and the air and to protect the front lens elements from the organic solvent. The light sheet is generated digitally\cite{Keller2008a,Keller2008b} by scanning the excitation beam across the focal plane of the detection objective. This generates incoherent illumination resulting in fewer artefacts.%cite{StelzerNatMeth2014}.
Additionally each line is specimen is illuminated with the same intensity creating a homogeneous light sheet which is particularly advantageous for the quantitative investigation of large samples in their entirety.

\subsection{Detection unit}

In it's simplest form, the fluorescence detection unit of a light sheet microscope is a conventional wide-field microscope: an objective lens, a fluorescent filter and a tube lens for an image on a wide-field detector. Fluorescence is collected with an objective that is specifically designed for immersion in clearing solutions. The objective is equipped with a correction collar allowing for immersion in media with refractive indices ranging from 1.41 to 1.52 (Olympus, XLSLPLN25XGMP). To image the whole volume of interest the objective has a relatively low magnification of 25x and a large working distance of 8mm yet a high numerical aperture (1.0NA) affords high resolution imaging. A tube lens of 200mm creates the primary image on a sCMOS camera (Orca Flash4.0, Hamamatsu) with a chip of over 4 megapixels. In contrast to CCD sensors each pixel of a sCMOS camera sensor has a separate photodetector and amplification unit which means that sCMOS pixels can be read out independently from each other. The cell size of the OrcaFlash is $p^2= 6.5\micro m$ x $6.5\micro m$ over an active area of 13.3mm. With a 200mm tube lens the field of view in the sample is $480\micro m$.

In the rolling shutter data acquisition mode only a subset of adjacent horizontal pixel lines is simultaneously exposed and this active detection region is moved across the image sensor. With the delay between the exposure of two adjacent lines set to the minimum time required to read out a single line ($t=9.7\micro s$, see Table\ref{tab:resolution}) the entire frame of 2048 horizontal lines is activated within 19.86ms. Setting the line exposure time to $l=3ms$ results in approx. $k=\nicefrac{l}{t}\approx300$  lines being simultaneously exposed at any time. This detection area is swept from the top to the bottom of the chip and acts like a moving, virtual confocal slit corresponding to a width of $d_s=\nicefrac{k p}{M_{\text{eff}}}= 0.72m$ in the sample space. With the given values of line exposure and line read out time, the camera acquires images at a frame rate of $\nu = \nicefrac{1}{(2048t+l)} \approx 44Hz$.

@@

\begin{figure}
   \begin{center}
   \begin{tabular}{c}
   \includegraphics[width=\textwidth]{frame2.eps}
   \end{tabular}
   \end{center}
   \caption{\label{fig:frame2} Schematic of light sheet microscopy. (A) Fluorescence excitation (along x axis) and detection (along z axis) are operated on independent, perpendicular light paths where the excitation light sheet and the detection focal plane overlap. (B) The custom-made sample chamber is assembled with silicone bellows and seal rings. The clarified, fluorescently-labelled brain is mounted on a Teflon cylinder in the centre of the watertight chamber (C) and can be translated and rotated freely with piezo motors (D). (E) Mounting of a clarified, fluorescently labelled brain. The brain is glued onto a coverslip and inserted into an adapter that slides into the Teflon cylinder.} 
   \end{figure}

For the acquisition of whole brain data sets, we fix a clarified and fluorescently-labelled mouse brain with super glue to a coverslip (Figure\ref{fig:frame2},D). The brain is oriented along its axial orientation with the olfactory bulbs at the top and the cerebellum at the bottom of the coverslip. The coverslip is slid into a bottom adaptor and tightened with a plastic-capped grub screw. The bottom adaptor is inserted into a Teflon cylinder with the coverslip being positioned on the far side of the detection objective. Three different slits in the adapter correspond to varying distances to the detection objective and give variability in sample thickness for example to allow also for the mounting of rat brain hemispheres.

We designed a cubic, water-tight sample chamber (Figure\ref{fig:frame2},A) that allows access from all six sides while maintaining the 3D integrity of large, clarified and fluorescently-labelled mouse brains. The sample chamber is tightly bolted  to the optical breadboard while soft connections using silicone bellows allow for adjustive movements of the objectives and free 3D motion of the motor stages (Figure\ref{fig:frame2},B). All connections are sealed with rubber rings and silicone caulk and additionally tightened with cable binders. In this way the objectives can be refocused and realigned without compromising the watertight seal of the chamber (Figure\ref{fig:frame2},C). The clarified brains are imaged immersed in clearing solution composed of 63\% TDE in Phosphate buffered saline (PBS) and a refractive index of 1.45. To fill the entire volume of the sample chamber approximately 400ml clearing solution are needed. A motorized x-, y-, z-, $\theta$-stage (Physik Instrumente, see Table\ref{tab:optomechanics}) allows free 3D motion and rotation of the Teflon cylinder which reaches into the centre of the chamber (Figure\ref{fig:frame2},D). Illumination and detection axes are horizontal and the sampler rotation occurs around the vertical axis.

@@

\section{Microscope management software}

	\begin{figure}
   \begin{center}
   \begin{tabular}{c}
   \includegraphics[width=\textwidth]{connectivity.eps}
   \end{tabular}
   \end{center}
   \caption{\label{fig:connectivity} Schematic views of the hardware (A) and software (B) architectures of the microscope, showing physical and logical connections, respectively. The software architecture scheme clarifies the hierachic structure of the Murmex components. AO analog output, DO digital output, DI digital input.} 
   \end{figure}

\subsection{General design principles}

The many components of the microscope (camera, lasers, AOTF, stages, etc.) are orchestrated together and operated through a custom-made software. Since the microscope should work mainly without human intervention for many hours (a whole-brain tomography can last a couple of days) without losing proper synchronization, software design is not a secondary issue when building a light sheet microscope for whole-brain imaging. The overall architecture of the software and its practical implementation were chosen to satisfy several principles:

\begin{itemize}
\item{\emph{User-friendliness}. The light sheet microscope here presented was designed and developed by optical technology specialists yet needs to be operational in a multi-user environment. The greatest challenge therefore consists in making the use of the microscope easy and intuitive such that no formidable expertise is needed when it is operated by researchers with no expertise in optics and/or software development.}
\item{\emph{Working in a distributed environment}. The apparatus should deal with many independent hardware components, and with several data streams (see below). Individual processing steps and/or data transfer can be computationally intensive, and may need to be distributed over several computers to perform properly at run-time.}
\item{\emph{Robustness and ease of maintenance}. Any kind of hardware problem, from timing delay to real failure, must be automatically handled to avoid loss of precious data. Furthermore, developers must be able to fix existing code or to develop new one without having to change the whole software.}
\item{\emph{Reusability and scalability}. When a second or third microscope with basic similarity yet slightly specialised equipment is being designed in the same or in other laboratories, re-writing all the code from scratch is pure loss of time. Writing well structured, scalable and easily maintainable scientific software can help avoiding redundancy and boosting up productivity.}
\end{itemize}

Our microscope management software has been written in LabVIEW (National Instruments, Austin, US), as this language is particularly well suited for hardware control and data acquisition through pre-developed methods. An object-oriented programming framework \cite{castagna1997object} has been used, dividing the entire project into well-defined self-contained modules. This framework, besides helping defining the proper scope of each software component, guarantees robustness, ease of maintenance, scalability and reusability. Code is stored on GitHub (www.github.org) to help multiple users developing software simultaneously. The graphical user interface (GUI) provides intuitive control of the apparatus as well as deep personalization. In general, the intrument parameters are already set to optimal values stored on disk, and on-line tips on how to set or change the various settings are provided. Both GUIs and code are developed according to the LabVIEW Style Guide \cite{LabviewStyle}.

\subsection{Murmex}

	\begin{figure}
   \begin{center}
   \begin{tabular}{c}
   \includegraphics[width=\textwidth]{murmex.eps}
   \end{tabular}
   \end{center}
   \caption{\label{fig:murmex} This graph needs a caption.} 
   \end{figure}

To let the software operate efficiently and synchronously on different machines, the \href{http://sine.ni.com/nips/cds/view/p/lang/en/nid/212895}{Murmex} software development kit was used. This kit integrates the object-oriented programming paradigm with a messaging system, and can be used to create communicating yet independent software modules within a standardized programming scheme.

Basically, Murmex creates distributed finite state machines. These are design patterns in which one component proceed from one particular state to another based on the message it receives either from itself or from other components. For sending and receiving messages, Murmex uses the library \href{http://sine.ni.com/nips/cds/view/p/lang/en/nid/211065}{LabbitMQ}, which is a wrapper of \href{http://www.rabbitmq.com/}{RabbitMQ}, a message-oriented middleware facilitating message shipment and receival. These layers of abstraction are mostly hidden for the developers. The RabbitMQ server (message broker) needs to be installed either locally or somewhere in the network. The message-based communication allows for buffered, asynchronous and reliable message communication between several software components. Messages are routed by the broker to the specific software component by its ID. The LabbitMQ wrapper allows use of RabbitMQ inside the LabVIEW programming environment.

Murmex design pattern has twelve pre-defined common states (Initialise, Configure, Start, Stop, Proceed, Update, Reset, Inspect, Inform, Acknowledge, Fetch, Terminate) and a generic state for implementing custom states. Usually, just a few of this states are actually used in standard application. An example of a standard implementation of the Murmex state machine for a scanning system is shown in Fig. BLABLA. Each Murmex component streams information about its current status to its 'observers', other components specified in its configuration file. Using this oberserver/observee hierarchy, it is possible to design large and complex network of different components, running on whatsoever number of computers connected through a local connection or the Internet. A special component, named ServiceManager (also based on Murmex), which must be running on each computer, can start executables remotely. This allows the user to start up a sophisticated configuration with multiple components executed on different computers, but steered from one computer.

\subsection{System timing and software organization}

A schematic overview of the hardware and software architecture of our light sheet microscope is shown in Fig. BLABLA, respectively. One personal computer (Precision T5600, Dell, Round Rock, USA) controls all the remote hardware components with the exception of the camera, which is controlled by a second, dedicated computer (Precision T7500, Dell, Round Rock, USA) to handle and manage the enormous streams of data produced (see next section). Various instrumentations are addressed either through computer ports (as USB, RS232 and CameraLink) or via analog and/or digital signals generated by a DAQ board (NI PCIe-6353, National Instruments, Austin, USA).

Synchronization of the camera rolling shutter to the scanning of the light sheet is achieved by letting the camera acquiring in free-run mode, and using a trigger output from the camera itself to time the generation of the waveforms sent to the galvo drivers. In particular, the HSYNC signal from the camera is used, providing a true-type-logic (TTL) pulse everytime the rolling shutter steps from one line to the next one. This trigger signal is used as a clock by the DAQ board, so that the line producing the light sheet moves step by step with the rolling shutter. The same HSYNC signal is used to trigger the DAQ digital signal sent to the electro-optic laser modulator. In this case, since at each position of the scanning line the modulator should go trough at least two states (vertical polarization and horizontal polarization), HSYNC is used as a trigger while the clock used is the (much faster) internal clock of the DAQ board.

Each hardware component (AOTF, galvo scanners, electro-optical modulator, sample stages, camera, lasers) is managed through its own Murmex component. Since the galvo scanners and the electro-optical modulator are not addressed directly, but via analog and digital signals generated by the DAQ board, their software component just send analog and digital waveforms to an additional intermediate component responsible for DAQ board management. Beyond hardware-related software, two additional Murmex components (named Stack and Tomo) orchestrate the operation of the entire system.

The stack component is used to acquire single image stacks along the $z$ direction, and to this aim send specific messages to coordinate the Camera and the Stages components. Stacks are collected by moving continuously the sample along $z$ while keeping the camera acquiring in free run. The translation speed is such that in the time needed to collect a single frame the specimen has moved by a $z$ step specified in the GUI. The Camera component uses low-level libraries from Hamamatsu to save the data, producing one single \emph{.cxd} file (which is a proprietary tiff-like format from Hamamatsu) for each image stack. The name of this file and the path to it is specified by the Stack component.

The tomo component is used to image the entire sample by collecting many adjacent, parallel, partially-overlapping stacks. To this aim it exchanges messages with the Stack and the Stages components. The user can specify the volume to be imaged either by inserting the maximum $x$, $y$ and $z$ coordinates of a parallelepiped, or by providing a text file with the list of stacks to be collected ($x$ and $y$ coordinates, starting and ending $z$s). The Tomo component also inform the experimenter about the state of the acquisition by sending an email when it's finished or if any trouble is experienced.

All the software described here can be freely downloaded from \href{https://github.com/marcelvanthoff/Giorgio}{GitHub}. Beacuse of its modular structure and of the extensive documentation, it can be easily integrated and adapted in other microscopy systems.

\section{Data Management}

The volume of a mouse brain have been estimated from MRI measurements \cite{Kovacevic2005} as $\simeq 0.5$ $\text{cm}^3$; the same Authors find out that a parallelepiped encompassing the whole brain would have a volume of $\simeq 0.9$ $\text{cm}^3$. Given the pixel size of the camera and the magnification of the system, and assuming a $z$ step of 2 $\mu$m, the voxel size results $0.234\times0.234\times2$ $\mu\text{m}^3$. This means that a mouse brain will be represented with about 4.6 TeraVoxels, while a parallelepiped encompassing the brain will result in 8.2 TeraVoxels. Since the camera produces 16 bit images with no compression, each raw voxel occupies 2 bytes of disk space, resulting in 8.3 TB for the whole brain "alone" and 14.8 TB for the parallelepiped (see Table XXXX). However, image stacks are acquired with a 10\% of linear superposition, which is used later for image stitching. This redundancy expands the size of raw data to 10.2 TB and 18.3 TB, respectively.

This unprecedented amount of data challenges to the traditional way we interact with our data. Having far passed the point of taking your data home with you on a hard drive at the end of the experiment, data management in light sheet microscopy requires the development of a robust data management infrastructure, and of novel software tools to process the images in order to extract meaningful information.

\subsection{Data handling and storage}

The camera used in our light sheet microscope (Hamamatsu Orca Flash 4.0 v2) flushes the datastream directly to a dedicated workstation with 12 solid-state drives (SSD) operating in RAID 0 configuration, resulting in a virtual drive of 10.9 TB. Although this is already a large and expensive system for the current SSD technology, it is way smaller than the size of the raw datasets produced in each tomography. We thus connected the workstation to a larger network area storage (NAS) via a 10 Gbit/s network copper cable. The NAS operates in a safer data modality (RAID 5) with 32.4 TB of useful disk space. The NAS is then connected by a dedicated 10 Gbit/s link provided by the Consortium GARR (the Public Institution managing research networks in Italy) to CINECA (the main italian supercomputing center). The data flow scheme is summarized in Figure \ref{fig:DataFlow}.
